### 2023-12-14 manuscriptlongversion.Rmd
Comment on manuscript

---
title: "Trauma treatment delay"
author: "Anton Wasielewski"
subtitle: A retrospective cohort study on the factors affecting the delays in trauma
  care
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
editor_options:
  markdown:
    wrap: 72
---

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(gtsummary)
library(gt)
library(knitr)
```

# Abstract

*Introduction:* Trauma remains a global health challenge, leading to
millions of deaths each year. The biggest impact can be seen in low to
middle income countries, but in high income countries, trauma systems
have been developed to lower the mortality and morbidity associated with
trauma. To further improve on these systems, conferences are held to
better the care patients receive and to identify errors made, called
<!-- Consider clarifying that delayed treatment may be one of the most 
common OFI-->
opportunities for improvement, e.g. delayed treatment. *Aim:* This
retrospective cohort study aims to identify patient level factors that
<!-- I would use "associated" rather than "linked"-->
are linked with receiving delayed treatment among trauma patients.
<!-- The first sentence is a bit complicated. Try to simplify, break 
into multiple sentences -->
*Material and Methods:* This study uses the database of Karolinska
university hospital, the equivalent of a level 1 trauma center, to run a
logistical regression linking the outcome, delayed treatment, to
variables, observed patient level factors. Patients treated for trauma
at the hospital between 2012 and 2022 were included, with ISS\>9 and/or
trauma code activation. Exclusion criterias were not being screened for
OFI and missing values in any of the variables, age, gender, ISS, SBP, RR,
GCS and number of injuries. *Results:* Higher ISS and RR was linked to
higher rates of delay in treatment (ISS, OR 1.03, 95% CI 1.01-1.06)(RR,
OR 1.04, 95% CI 1.01-1.08), while a lower GCS also was linked to delayed
treatment (OR 0.93, 95% CI 0.87-1.00). Number of injuries showed a
correlation to higher rates of delay, but not when adjusted for. The
remaining variables showed no correlation. *Conclusion:* Our study
suggests that it is mostly more severely injured and less awake patients
that receive delayed treatment.

*Inledning:* Trauma är fortfarande en global hälsoutmaning som skördar
miljontals liv varje år. Den största effekten kan ses i låg till
medelinkomstländer, men i höginkomstländer har traumasystem utvecklats
för att sänka morbiditeten och mortaliteten relaterat till trauma. För
att ytterliggare förbättra dessa system hålls konferenser för att
förbättra vården patienter får och identifiera misstag som gjorts, så
kallade förbättringsmöjligheter, t.ex. försenad vård. *Syfte:* Denna
retrospektiva kohortstudie syftar till att identifiera patientfaktorer
kopplade till försenad vård bland traumapatienter. *Material och Metod:*
Denna studie använder Karolinska universitetssjukhusets databas,
motsvarande ett traumacenter på nivå 1, för att köra en logistisk
regression som kopplar utfallet, försenad vård, till variabler,
observerade patientnivåfaktorer. Patienter som behandlades för trauma på
sjukhuset mellan 2012 och 2022 inkluderades, med ISS\>9 och/eller
traumakodaktivering. Exklusionskriterier var om patient inte screenades
för ofi samt saknade värden i de använda variablerna, ålder, kön, ISS,
SBT, AF, GCS och antal skador. *Resultat:* Högre ISS och AF var kopplade
till högre grad av försenad vård (ISS, OR 1.03, 95% CI 1.01-1.06)(RR, OR
1.04, 95% CI 1.01-1.08), medan ett lägre GCS också var kopplad till
försenad vård (OR 0.93, 95% CI 0.87-1.00). Antalet skador visade först
ett samband med högre förseningsfrekvens, men inte i den justerade
analysen. De återstående variablerna visade ingen korrelation.
*Slutsats*: Vår studie tyder på att det mestadels är svårare skadade och
mindre vakna patienter som får försenad behandling.

# Introduction

## Definition

Trauma, defined as the clinical entity composed of physical injury and
the body's associated response [@gerdin_risk_2015], is and has long been
a major cause of death around the world [@WHO_2021]. 
<!-- The sentence below is long and overly complicated. Are you saying 
that trauma systems is an important part of healthcare and its implementation
has reduced mortality and morbidity?-->
Within modern
hospital care, the trauma system stands as an important part in
healthcare and is crucial to lowering mortality and morbidity in injured
patients through many means, such as hospital care, patient follow up
and prevention programs [@Choi_2021].

## Trauma Statistics
<!-- I suggest you name this section something else, like "Trauma globally" 
or something like that. Also, divide the second paragraph into two or 
three paragraphs-->
Trauma is often divided into two subgroups, blunt force trauma and
penetrating trauma. Blunt force trauma is when an object or force
strikes the body, often causing bruising, broken bones or deep cuts.
Examples of blunt trauma could be car crashes, falls or direct blows to
the body. Penetrating trauma is when an object pierces the skin or body
and creates an open wound. Examples of penetrating trauma are gunshot
wounds and stab wounds [@NIGMS_2020]. Differences exist between the two
groups, as blunt trauma patients tend to be more injured on arrival to
the hospital and as such, require more resources and are hospitalized
for a longer periods [@FITCH2019].

Today, trauma takes the lives of around 4,4 million people each year,
almost 8% of all deaths [@WHO_2021]. In the United States, trauma is the
4th leading cause of death among the general population and the leading
cause for people between the ages of 1 and 44 [@AAST_2020]. It is also
an important cause of hospitalization and morbidity among all age
groups, including seniors, and is responsible for an estimated 10% of
all years lived with a disability globally. This has a significant
burden on social and economic level, costing countries billions of US
dollars each year in healthcare and lost productivity [@connolly_2018].
Studies estimate the cost of trauma care being between 18,500\$ and
41,500\$ per patient in high income countries (HIC), depending on the
country [@Willenberg_2012]. However, it has been shown that this burden
is not evenly distributed between or within countries. Many social
factors such as age, sex and social status play a major role in the risk
of dying from trauma, with young men with low socio-economic status
being at most risk [@WHO_2021]. But it is not only patient level factors
that affect what effect trauma has on people. About 90% of all trauma
related deaths occur in low- and middle-income countries (LMIC), with
death rates by trauma also being higher than in HIC. Even within these
LMIC, people of poorer socio-economic status have higher death rates
from trauma. Problems identified within these countries that contribute
to these statisitics were infrastructure, education and training,
attributed to lack of funding, brain drain to HIC and lack of
availability of basic amenities [@Shanthakumar_2021]. In HIC where
funding and governance over the healthcare system is better designed,
functioning trauma systems and dedicated trauma centers exist that have
been shown to lower mortality but also improve functional outcome in
trauma patients [@Nirula_2006].

## Trauma systems
<!-- Split long paragraphs -->
Trauma systems are infrastructures that exist to provide and optimize
care for injured patients starting with injury recognition and triage,
transport to appropriate trauma center, inpatient care and outpatient
follow-up. Beyond the clinical side, trauma systems work with outreach,
education and advocacy, data collection through registries, research,
funding, and disaster preparedness and response. A comprehensive and
functioning trauma system requires strong leadership and engagement at
the trauma center, regional and national level [@connolly_2018]. This
system is crucial to provide care for trauma patients, both in reducing
morbidity and mortality in this patient group. Earlier studies from
Sweden have shown that treating severely injured patients at a trauma
center is associated with a 41% lower adjusted 30-day mortality rate
compared to being treated at a non trauma care center due to them being
more capable of treating these patients, with potential survival benefit
increasing with higher injury severity [@Candefjord_2020]. Other studies
have shown similar results, showing that treating injured patients at
trauma centers is associated with a 15-30% decrease in mortality
[@Moran_2018; @Celso_2006; @MacKenzie_2006; @Haas_2012]. During the
last three decades, the introduction of trauma systems has contributed
to lowering the incidence of preventable death. This is attributed to
the improvements in care for acute brain injuries and bleeding control.
The incidence of late death because of sepsis and multiple organ failure
has also been lowered, possibly a result of better and earlier
resuscitation [@Asensio_2008]. Functioning trauma systems saves lives in
the hospital, but their work outside of the hospital is just as
important. Data collection from injured patients, such as mechanism of
injury or mortality, are essential for creating databases that can be
used for research. In turn, that research can be used for planning
injury prevention programs that target the most common injuries in the
most efficient ways, e.g. teen drivers, children, specific occupations
etc. These injury prevention programs can be planned on trauma center,
organizational or government level[@ACS].

Trauma teams are multidisciplinary and operate in these centers. They
play a pivotal role in the treatment of the trauma patient, as they
provide the initial care in the critical stage of trauma. In Sweden,
trauma teams are lead by a team leader who is a surgeon, and include
practitioners from the specialities of intensive care, orthopedics,
nursing and support staff[@västra_G]. For the trauma team to be
mobilized, a trauma code has to be activated, often by an emergency nurse.
The nurse uses information gathered by the first responders to assess if
the patient fits any of the criteria for trauma code. Most healthcare
facilities have established criteria or guidelines that trigger trauma
code activation. These criteria typically include mechanisms of injury,
physiological criteria, anatomical criteria and other specific
indicators of severe trauma. There are different levels of trauma code
with different criteria correlating with severity of injury, with level
1 mobilizing the most personnel to the trauma room. Once in the trauma
room, the team works systematically to manage the patients injuries.
They handle the most urgent problems first, such as airways and
breathing, with the aim to rapidly assess and stabilise the patient,
prioritise their injuries and arrange for site of definitive care.
[@Hedberg_2020]

## Mortality and morbidity (MoM) conferences
<!-- Consider renaming this section as "Opportunities for improvement" -->
Many hospitals have a trauma registry where they log the patients
information and timeline of what happened. The extent of the registry
varies, with HIC having more complete registries. In Sweden, 48 out of
the 49 hospitals receiving trauma patients are connected to the national
registry SweTrau, which functions as a national database designed to
enable scientific processing with epidemiological technology
[@SweTrau_2021] This logging of information is important work, as a
cornerstone of trauma quality improvement programs is multidisciplinary
MoM conferences. The MoM conference is a meeting where different
specialities who work with trauma care sit down and discuss deaths and
complications in order to look for preventable factors. These
conferences are performed in many hospitals globally, almost everywhere
where there are formal medical specialty departments and sometimes in
smaller hospitals as well [@WHO_2009]. In Sweden, many hospitals have
mortality conferences on deceased patient cases, but only one has MoM
conferences to better their trauma care on patients who lived
[@SweTrau_2023]. The reason being that this expansion of MoM conferences
takes time and resources, which is why all hospitals don't implement
this.

The endpoint of these MoM conferences are opportunities for improvement
(OFI). At the end of the conference, consensus is reached regarding the
existence of an OFI and implementation of corrective action. This
process is effective, as it has been shown that this review is
associated with high-quality trauma care [@Santana_2014]. Examples of
OFI may include lack of resources and management errors. One common OFI
is delayed treatment. According to previous studies, among the
preventable or possibly preventable deaths in trauma patients, delay in
treatment has been identified as a major error contributing to death,
found in up to 52.9% of patients in said group [@Teixeira_2007]. Delayed
treatment has been shown to have adverse effects on patients, showing
why it is such an important issue and the need to develop strategies to
combat [@Sampalis_1995; @Sampalis_1994]. Although it is such a common
OFI, the patient level factors associated with delayed treatment are
poorly understood. There may be several factors that correlate with
receiving delayed treatment, but they have yet to be identified. Mapping
these factors may help in identifying patients that might be at risk for
receiving delayed treatment before it happens.

## Aim
<!-- Make sure the aim here is exactly the same as in the abstract. Also,
I suggest that your aim is to determine the associations between different 
patient level factors and receiving delayed treatment -->
The aim of this paper is to determine what different patient level
factors affect the risk of receiving delayed treatment at a trauma
center.

# Methods

## Setting

The Karolinska University hospital, which is the equivalent of a trauma
level 1 hospital, admits 1500 trauma patients each year. To be included in
 the trauma registry, a patient must be over 15
years of age, had ISS\>9 and/or had trauma code activation. They get
added to both the Karolinska Trauma registry as well as the national
trauma registry (SweTrau). The registry includes data on vital signs,
injuries, interventions and patient information. Another database, the
trauma care quality database includes information relevant for MoM
conferences. At Karolinska University Hospital, patients who die during
hospital stay gets their case sent to a MoM everytime. However, for the
rest of the patients, two specialized nurses determine whether they are.
Several audit filters are set up to flag for patients with possible
OFIs. Audit filters may include low SBP, high RR, EMS scene time greater
than 20 mins etc. Flagged patients are then reviewed by these two nurses
who decide if there is a possible OFI, in which case their case is sent
to a MoM conference for further review.

## Study design

We conducted a registry based cohort study using data from the trauma
registry and trauma care quality database at the Karolinska University
Hospital in Solna. The trauma registry includes 14,022 patients treated
between 2012 and 2022. The trauma care quality database is a subset of
the trauma registry and includes about 8000 patients selected for review
between 2013 and 2022. This project linked the two databases and
assessed how different patient level factors, such as age, sex, blood
pressure, and injury severity, are associated with delayed treatment
using logistic regression.

## Outcome

The outcome is the OFI delayed treatment. An OFI is any failure in care,
e.g. clinical judgement error, any preventable or potentially
preventable death, missed diagnosis, technical error and delayed
treatment as identified and decided my the MoM conference. The outcomes
we used were "long time until CT" and "long time until operation",
which will be referred to as delayed treatment from now on.

## Participants

The database only includes patients 15 years old and above, ISS \>9
and/or trauma team activation. Inclusion in the study further required
that the patients has been assessed for OFIs. Patients were excluded if
they were missing information in one of the examined patient factors.

## Variables

The independent variables are gender, GCS on arrival, respiratory rate
on arrival, systolic blood pressure on arrival, number of injuries, ISS
and age. Of these variables, GCS, respiratory rate, systolic blood
pressure, number of injuries, ISS and age are used as continuous
variables in the analysis. Gender is used as a categorical variable.

For missing hospital values in GCS, RR and SBP, values collected from
EMT personnel before arrival to hospital were used. This was
accomplished by identifying missing data in the column for the hospital
value of the variable and, if available, replacing the row in the
dataset with the prehospital value. If a prehospital value is missing,
the patient is excluded from the study.

## Data sources/measurement

The data comes from the trauma registry and trauma care quality
database.

## Bias

Selection bias as this is a retrospective post hoc cohort study where
the outcome has occurred before the study began.

Misclassification bias as the conference may reach a consensus that is
wrong, alternatively that the two specialized nurses going through the
patients flagged for audit filters might misclassify patients or that
the audit filter misses a patient with ofi.

## Study size

The study size is all of the eligible patients that was treated at the
hospital between 2012 and 2022. Starting 2013 only a subgroup of
patients were screened for OFIs, but starting 2017 all patients that are
included in the database are screened for OFIs. Out of the 14,022
patients in the database, 7,378 were included in the study, with the
rest being excluded due to either not being screened for ofi or missing
values in the examined variables. 72 patients were deemed to have
received delayed treatment. However, only 70 patients with the
registered outcome were included in the study due to missing values
among two of the patients.

## Statistical methods

Multivariable and univariable logistical regression. A 5% significance
level and 95% confidence levels will be used.

Rstudio was used to clean and prepare the data from the database. The
data was processed to have the variables customized to fit into the
logistical regression model, such as making them into factors, renaming
outcomes in the database and replacing or removing unknown values. The
variables were either binary or continuous. While writing the code, it
was tested on scrambled data so to avoid selection bias as much as
possible. When the code was finished, it was run on the unscrambled data
for a result.

Using the one in ten rule, 7 independent variables were used for the
logistical regression given that we have 70 patients with delayed
treatment. The one in ten rule states that for every variable examined,
10 events are needed in the group. In this case, this means that because
we have 70 patients with the examined outcome, we can have 7 variables.

The logistical regression is run through Rstudio using the glm command.
It was ran several times to make an adjusted table and 7 unadjusted, one
for each variable. To make the tables, the package gtsummary was used to
visualize the adjusted and unadjusted regression individually, then
merging them together into one table.

This process was the only time generative AI was used in the writing process. It
was used to make snippets of code for a specific purpose, revise and
optimize code and find solutions when written code was not functioning
properly.

## Ethical considerations

Ethical permit is required and exists. The ethical review number is
2021-02541 and 2021-03531. An ethical problem that might be brought up
is the anonymity of the patients. In the registry, patients social
security number is present and could theoretically be used to identify
the patient in question. This will not be a problem though, as in the
analysis, the social security number is removed from the data before the
logistical regression is ran. Furthermore, the results will be presented
with patients grouped together, not individually. This will make it
impossible for anyone to identify patients. During the analysis, a VPN
was used to access the database to ensure that no data was leaked or
used for other purposes.

# Results

Out of the 14,022 patients receiving trauma care at Karolinska
University Hospital, 5,710 patients were not screened for ofi, which
excludes them from the study and leaves 8,312 patients that have been
screened for ofi.

As we were running the logistic regression, additional patients were excluded from
the study due to missing data. Among the 8,312 patients screened for
ofi, 7,378 were included in the study. 934 patients were excluded from
the study due to missing values in one or more variables. This left us
with 70 patients in the delayed treatment group and 7,308 patients in
the no delay to treatment group. Figure 1 below illustrates the
exclusion and inclusion of patients.

![](Rplot.png)

<!-- You need to refer to tables and figures by numbers in the text, 
i.e. Figure 1, Table 1 and so on -->
Table 1 shows the characteristics of the group used in the logistic
regression. Most of the patients were men, 68%, with that being a bit
higher in the delayed treatment group, 74%. The mean value for the ISS
score was 10.18, with that value being higher in the outcome group, at
16.81. The mean age of the patients was 45.18 years old, with it being
around the same in the delayed treatment group at 45.50. The mean number
of injuries among the delayed treatment patients was at 6.33, which is
higher than in the overall group at 4.76. The mean SBP was 137.33, with
a standard deviation of 25.69. Mean respiratory rate was 18.59 with it
being slightly higher in the outcome population at 20.39. Lastly, mean
GCS was 13.93, being lower among patients with delayed treatment at
12.56.

```{r message = FALSE}
table1
```

In the following table<!-- See comments above-->, the results from the logistical regression
analysis is displayed.

The odds ratio is calculated both unadjusted and adjusted for other
variables<!-- What does "other" mean here?-->, as well as a 95% confidence interval. 
The p value is also shown.

<!-- Use "associated" instead of correlated or linked etc-->
The analysis shows that some variables seem to have a correlation with
patients receiving delayed treatment. ISS and respiratory rate (RR) both have
a trend that patient with lower values seem to have a lower incidence of
delayed treatment, in both unadjusted (ISS, OR 1.04, 95% CI 1.02-1.05)(RR,
OR 1.06, 95% CI 1.02-1.09) and adjusted analyses (ISS, OR 1.03, 95% CI
1.01-1.06)(RR, OR 1.04, 95% CI 1.01-1.08). GCS had the opposite
correlation, with higher GCS often having lower incidence of delayed
treatment, both unadjusted (OR 0.88, 95% CI 0.83-0.94) and adjusted (OR
0.93, 95% CI 0.87-1.00). Number of injuries had the correlation of
higher number of injuries having higher incidence of delayed treatment
unadjusted (OR 1.06, 95% CI 1.02-1.10), but that correlation was not
seen when adjusted for. The three remaining variables, age, gender and
SBP, showed no correlation with the observed outcome.

```{r message = FALSE}
result1
```

# Discussion

These results show a possible correlation between some variables and the
presence of delayed treatment. A lower GCS had a correlation with
delayed treatment while a higher ISS had the same correlation. This
suggests that a more heavily injured and less awake patient is more
prone to receiving delayed treatment. Compared to a previous
study analysing the same data but with OFI as the outcome, this result
is similar to what they found using the same database [@Pat_factors]. As
the mentioned study suggests, this could be explained by that lower GCS as well as higher ISS
indicates severe trauma. More injured patients
often requires a more complex management and more interventions.
More interventions leave more room for delay to these interventions, why
they are more prone to delay in treatment. Respiratory rate also had the
correlation that lower respiratory rate correlated with lower rates of
delayed treatment among the patients. As the interval is \>10 to \<29 in
the triage guideline, both too low and too high respiratory rate is not
desirable[@EHCP2015]. This might have a similar explanation to ISS and
GCS, that patients with tachypnea might receive higher rates of delayed
treatment because they are more severely injured, therefore a higher
risk of missing a possible injury. The previous study also found this, but
the correlation dissapeared when adjusted for, which was hypothesised
being due to its shared variability with the ISS
variable[@Pat_factors]. In this study, this was the case for number of
injuries, as it and ISS is often correlated as a more severely injured
patient will both score higher on ISS and number of injuries. Therefore,
we hypothesize that the effect of number of injuries become
insignificant when adjusting for ISS.

Not many studies regarding trauma and patient level factors have had
delayed treatment as their outcome, while many analysing mortality. Some
similarities exist with which patient level factors predict the outcome,
with higher mortality being seen in patients with lower GCS and higher
ISS [@Mort_sam][@Mort_hef]<!-- Are these two references?-->. Increased age was also a factor linked to
mortality, which was not seen to be linked to delayed treatment in this
study. An older patient is more likely to have co-morbidities that might
affect their mortality but not increase their likelihood of receiving
delayed treatment the same way ISS and GCS does[@Gioffre2018].

When looking at the patient demographics, it is clear that the majority
admitted to trauma care are men, with it being slightly higher in the
delay to treatment group. However, not high enough to be statistically
significant. The fact that gender did not show an increased odds ratio
should be seen as a positive sign. It suggests that among trauma
patients, there is no systematic delay based on this patient level
factor.

#### Limitations

In this study, one of the limitations was the number of people having
the recorded outcome being quite small, leading us into having to limit
the number of variables we wanted to examine. Furthermore, the original
plan was to divide RR, SBP and GCS into revised trauma score (RTS)
categories, from 1-4. However, many of the RTS variables had few to no
patients with the observed outcome. This led to separation in the
logistic regression in these variables, referring to when an outcome
variable separates a predictor variable completely, which occurs when
the predictor is only associated with one outcome value [@OARC]. In the
case of these variables, it was only RTS 3 and 4 who had patients with
the observed outcome, which left RTS 1 and 2 with confidence intervals
of 0.00 to infinity. This way would have been the best way to examine
these variables, but as that was not possible with this group of
patients, we resorted to making these variables continuous. We
considered expanding the outcome group to include more OFI, but opted
for the solution used in the study instead as we were solely interested
in this OFI for this study.

The problem mentioned above with too few patients in the outcome group
made this into a post hoc analysis. As we did not know the number of
people who received delayed treatment when we handled the scrambled
data, we ran the code on the real data when we had too many variables
for the dataset. Therefore, we had to change the variables after having
seen the real data, making it a post hoc analysis. This could in theory
open up for possible selection bias, as we might choose variables which
give us results we are looking for. However, all the variables included
in the final analysis were already included in the first one. Variables
were only removed, not added after having seen the real data. The best
way to have gone about this would have been to just look at the amount
of people who received delayed treatment before choosing the variables.
Then, choosing an amount of variables that would work with the amount of
patients. This would have avoided making it post hoc and changing the
variables after seeing the data.

On discussing possible biases, misclassification bias might also be
<!-- Do not use contractions, i.e. can't, doesn't etc -->
present in this study, though it is something that we cannot control in
this study. As previously mentioned, the possible existence of an OFI is
decided during the MoM conferences, where they then add it to the
database. However, they might not always be right in their judgement and
misclassify from time to time, leading to patients landing in the wrong
group in the analysis. Also, some patients might have been missed even
before reaching the MoM conference. The patients included in the
conference are selected through audit filters and two specialized
nurses, and patients might be missed in this stage. This bias should not
be big enough to have a significant impact on the results, but can be
kept in mind.

<!-- Not so sure about this section. What do you mean that the logistic
regression assumes that the data is binomial? Do you mean that it requires 
a binary outcome? I think you discuss potential limitations well enough in 
the previous paragraphs, so you don't really need this -->
Another limitation with this study is that logistic regressions
assumes that the data is binomial, which the data used is not. The
analysis assumes a linear correlation between the dependent variable and
the independent variable, which is best analysed in a normal
distribution. However, Fischers test showed a significant difference
between the groups in 4 of the 7 variables, all of them being the ones
who showed an increased or decreased risk of delayed treatment, showing
that these results still are significant.

#### Clinical applications

<!-- Revise this section now after our meeting -->

Further analysing the results and it's clinical applications, the
results can be interpreted as that it is mostly injured and less awake
patients that receive delayed. The clinical applications for this are
not clear, but caregivers should already be aware that the more injuries
a patients has, the more room for error exists. It is good though to
receive a reminder that it's important to always work towards minimizing
the delay for these injured patients.

#### Generalisability

Discussing the generalisability of this study, we included the all the
trauma patients treated and Karolinska University Hospital and excluded
40.7% of that population based on that they had not been screened for
ofi. Additional patients were excluded because of missing data, but no
distinct groups except the not screened patients were removed. On that
basis, this population should be a general representation of trauma
patients at the equivalent of a level 1 trauma center, making this study
generalisable. This is in line with our aim, as it was to make it
applicable to the whole trauma patient population. However, as it is a
single center study, where its result only illustrates the situation in
Stockholm, it is less applicable nationally or globally. This does not
make it only applicable to the population of Stockholm, but is something
to keep in mind.

<!-- Think about what components of your study is generalisible and to 
where. I think that your results are generalisible to  a large extent, in 
the sense that severly injured patients are at higher risk of delayed 
treatment, also in other contexts -->

#### Future studies

In future studies building upon this one, various methods to further
examine the correlation between delayed treatment and various variables,
alternatively using another OFI as the outcome variable. Examining
different variables such as highest hospital care lever, intubation
status or first emergency procedure performed might be interesting
variables, among others. To be able to do this, more patients in the
outcome group would be needed, either by expanding the register to
several hospitals, or expanding the outcome group by adding different
ofi. This however would make the study less specific for delayed
treatment. Further studies might also look into different ofi other than
delayed treatment, such as missed injury, faulty communication or
triage.

## Conclusion

In conclusion, based on this small population, the results suggests that
<!-- I would say "with affected consciousness" instead of "less awake" -->
it is mostly more injured and less awake trauma patients that receive
delayed treatment. However, due to the limitations of this study, the
result should only be seen as something to keep in mind when treating
patients as further studies need to be conducted on this subject.

#