### 2024-01-09 manuscriptlongversion.Rmd
possible final draft

---
title: "Trauma treatment delay"
author: "Anton Wasielewski"
subtitle: A retrospective cohort study on the factors affecting the delays in trauma
  care
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: bibliography.bib
csl: vancouver.csl
editor_options:
  markdown:
    wrap: 72
---

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(gtsummary)
library(gt)
library(knitr)
```

# **Trauma and the patient level factors linked with delayed treatment**

Introduction: Trauma remains a global health challenge, leading to
millions of deaths each year. The biggest impact can be seen in low to
middle income countries, but in high income countries, trauma systems
have been developed to lower the mortality and morbidity associated with
trauma. To further improve on these systems, conferences are held to
better the care patients receive and to identify errors made, called
opportunities for improvement, one of the more common being delayed
treatment. Aim: This retrospective cohort study aimed to determine the
associations between different patient level factors and receiving
delayed treatment. Material and Methods: This study used the database of
Karolinska university hospital to run a logistic regression linking the
outcome, delayed treatment, to variables, observed patient level
factors. Patients treated for trauma at the hospital between 2012 and
2022 were included, with ISS\>9 and/or trauma code activation. Exclusion
criterias were not being screened for OFI and missing values in any of
the observed variables. Results: Higher ISS and RR was linked to higher
rates of delay in treatment (ISS, OR 1.03, 95% CI 1.01-1.06)(RR, OR
1.04, 95% CI 1.01-1.08), while a lower GCS also was linked to delayed
treatment (OR 0.93, 95% CI 0.87-1.00). Number of injuries showed a
correlation to higher rates of delay, but not when adjusted for. The
remaining variables showed no correlation. Conclusion: Our study
suggests that it is mostly more severely injured and less awake patients
that receive delayed treatment.

**Trauma och patientnivåfaktorerna kopplade till försenad behandling**

Inledning: Trauma är fortfarande en global hälsoutmaning som skördar
miljontals liv varje år. Den största effekten kan ses i låg till
medelinkomstländer, men i höginkomstländer har traumasystem utvecklats
för att minska morbiditeten och mortaliteten relaterat till trauma. För
att ytterligare förbättra dessa system hålls konferenser för att
förbättra vården patienter får och identifiera misstag som gjorts, så
kallade förbättringsmöjligheter, där försenad vård är en av de
vanligaste. Syfte: Denna retrospektiva kohortstudie syftade till att
identifiera associationen mellan patientfaktorer och försenad vård.
Material och Metod: Denna studie använde Karolinska
universitetssjukhusets databas för att köra en logistisk regression som
kopplar utfallet, försenad vård, till variabler, observerade
patientnivåfaktorer. Patienter som behandlades för trauma på sjukhuset
mellan 2012 och 2022 inkluderades, med ISS\>9 och/eller
traumakodaktivering. Exklusionskriterier var om patient inte screenades
för ofi samt om de saknade värden i de undersökta variablerna. Resultat:
Högre ISS och AF var kopplade till högre grad av försenad vård (ISS, OR
1.03, 95% CI 1.01-1.06)(RR, OR 1.04, 95% CI 1.01-1.08), medan ett lägre
GCS också var kopplad till försenad vård (OR 0.93, 95% CI 0.87-1.00).
Antalet skador visade först ett samband med högre förseningsfrekvens,
men inte i den justerade analysen. De återstående variablerna visade
ingen korrelation. Slutsats: Vår studie tyder på att det mestadels är
svårare skadade och mindre vakna patienter som får försenad behandling.

Keywords: Trauma, Delay to treatment, Patient factors.

# **Introduction**

## **Definition**

Trauma, defined as the clinical entity composed of physical injury and
the body's associated response (1), is and has long been a major cause
of death around the world (2). The trauma system plays an important part
in modern healthcare and has shown to lower mortality and morbidity in
injured patients through many means, such as hospital care, patient
follow up and prevention programs (3).

## **Trauma globally**

Trauma is often divided into two subgroups, blunt force trauma and
penetrating trauma. Blunt force trauma is when an object or force
strikes the body, often causing bruising, broken bones or deep cuts.
Examples of blunt trauma could be car crashes, falls or direct blows to
the body. Penetrating trauma is when an object pierces the skin or body
and creates an open wound. Examples of penetrating trauma are gunshot
wounds and stab wounds (4). Differences exist between the two groups, as
blunt trauma patients tend to be more injured on arrival to the hospital
and as such, require more resources and are hospitalised for a longer
period (5).

Today, trauma takes the lives of around 4,4 million people each year,
almost 8% of all deaths (2). In the United States, trauma is the 4th
leading cause of death among the general population and the leading
cause for people between the ages of 1 and 44 (6). It is also an
important cause of hospitalisation and morbidity among all age groups,
including seniors, and is responsible for an estimated 10% of all years
lived with a disability globally. This has a significant burden on
social and economic level, costing countries billions of US dollars each
year in healthcare and lost productivity (7). Studies estimate the cost
of trauma care being between 18,500\$ and 41,500\$ per patient in high
income countries (HIC), depending on the country (8).However, it has
been shown that this burden is not evenly distributed between or within
countries.

Many social factors such as age, sex and social status play a major role
in the risk of dying from trauma, with young men with low socio-economic
status being at most risk (2). But it is not only patient level factors
that affect what effect trauma has on people. About 90% of all trauma
related deaths occur in low- and middle-income countries (LMIC), with
death rates by trauma also being higher than in HIC. Even within these
LMIC, people of poorer socio-economic status have higher death rates
from trauma. Problems identified within these countries that contribute
to these statistics were infrastructure, education and training,
attributed to lack of funding, brain drain to HIC and lack of
availability of basic amenities (9). In HIC where funding and governance
over the healthcare system is better designed, functioning trauma
systems and dedicated trauma centers exist that have been shown to lower
mortality but also improve functional outcome in trauma patients (10).

## **Trauma systems**

Trauma systems are infrastructures that exist to provide and optimise
care for injured patients starting with injury recognition and triage,
transport to appropriate trauma center, inpatient care and outpatient
follow-up. Beyond the clinical side, trauma systems work with outreach,
education and advocacy, data collection through registries, research,
funding, and disaster preparedness and response. A comprehensive and
functioning trauma system requires strong leadership and engagement at
the trauma center, regional and national level (7). This system is
crucial to provide care for trauma patients, both in reducing morbidity
and mortality in this patient group. Earlier studies from Sweden have
shown that treating severely injured patients at a trauma center is
associated with a 41% lower adjusted 30-day mortality rate compared to
being treated at a non trauma care center due to them being more capable
of treating these patients, with potential survival benefit increasing
with higher injury severity (11). Other studies have shown similar
results, showing that treating injured patients at trauma centers is
associated with a 15-30% decrease in mortality (12--15).

During the last three decades, the introduction of trauma systems has
contributed to lowering the incidence of preventable death. This is
attributed to the improvements in care for acute brain injuries and
bleeding control. The incidence of late death because of sepsis and
multiple organ failure has also been lowered, possibly a result of
better and earlier resuscitation (16). Functioning trauma systems saves
lives in the hospital, but their work outside of the hospital is just as
important. Data collection from injured patients, such as mechanism of
injury or mortality, are essential for creating databases that can be
used for research. In turn, that research can be used for planning
injury prevention programs that target the most common injuries in the
most efficient ways, e.g. teen drivers, children, specific occupations
etc. These injury prevention programs can be planned on trauma center,
organisational or government level(17).

Trauma teams are multidisciplinary teams and operate in these centers.
They play a pivotal role in the treatment of the trauma patient, as they
provide the initial care in the critical stage of trauma. In Sweden,
trauma teams are led by a team leader who is a surgeon, and include
practitioners from the specialities of intensive care, orthopaedics,
nursing and support staff(18). For the trauma team to be mobilised, a
trauma code has to be activated, often by an emergency nurse. The nurse
uses information gathered by the first responders to assess if the
patient fits any of the criteria for trauma code. Most healthcare
facilities have established criteria or guidelines that trigger trauma
code activation. These criteria typically include mechanisms of injury,
physiological criteria, anatomical criteria and other specific
indicators of severe trauma. There are different levels of trauma code
with different criteria correlating with severity of injury, with level
1 mobilising the most personnel to the trauma room. Once in the trauma
room, the team works systematically to manage the patient's injuries.
They handle the most urgent problems first, such as airways and
breathing, with the aim to rapidly assess and stabilise the patient,
prioritise their injuries and arrange for a site of definitive care.
(19)

## **Opportunities for improvement**

Many hospitals have a trauma registry where they log the patients
information and timeline of what happened. The extent of the registry
varies, with HIC having more complete registries. In Sweden, 48 out of
the 49 hospitals receiving trauma patients are connected to the national
registry SweTrau, which functions as a national database designed to
enable scientific processing with epidemiological technology (20). This
logging of information is important work, as a cornerstone of trauma
quality improvement programs is multidisciplinary mortality and
morbidity (MoM) conferences. The MoM conference is a meeting where
different specialties who work within trauma care sit down and discuss
deaths and complications in order to look for preventable factors. These
conferences are performed in many hospitals globally, almost everywhere
where there are formal medical specialty departments and sometimes in
smaller hospitals as well (21). In Sweden, many hospitals have mortality
conferences on deceased patient cases, but few has MoM conferences to
better their trauma care on patients who lived (22). The reason being
that this expansion of MoM conferences takes time and resources, which
is why all hospitals don't implement this.

The endpoint of these MoM conferences are opportunities for improvement
(OFI). At the end of the conference, consensus is reached regarding the
existence of an OFI and implementation of corrective action. This
process is effective, as it has been shown that this review is
associated with high-quality trauma care (23). Examples of OFI may
include lack of resources and management errors. One common OFI is
delayed treatment. According to previous studies, among the preventable
or possibly preventable deaths in trauma patients, delay in treatment
has been identified as a major error contributing to death, found in up
to 52.9% of patients in said group (24). Delayed treatment has been
shown to have adverse effects on patients, showing why it is such an
important issue and the need to develop strategies to combat (25,26).
Although it is such a common OFI, the patient level factors associated
with delayed treatment are poorly understood. There may be several
factors that correlate with receiving delayed treatment, but they have
yet to be identified. Mapping these factors may help in identifying
patients that might be at risk for receiving delayed treatment before it
happens.

## **Aim**

This retrospective cohort study aimed to determine the associations
between different patient level factors and receiving delayed treatment.

# **Methods**

## **Study design**

We conducted a registry based cohort study using data from the trauma
registry and trauma care quality database at the Karolinska University
Hospital in Solna. The trauma registry includes 14,022 patients treated
between 2012 and 2022. The trauma care quality database is a subset of
the trauma registry and includes about 8000 patients selected for review
between 2013 and 2022. This project linked the two databases and
assessed how different patient level factors, such as age, sex, blood
pressure, and injury severity, are associated with delayed treatment
using logistic regression.

## **Setting**

The Karolinska University hospital, which is the equivalent of a trauma
level 1 hospital, admits 1500 trauma patients each year. To be included
in the trauma registry, a patient must be over 15 years of age, had
ISS\>9 and/or had trauma code activation. They are added to both the
Karolinska Trauma registry as well as the national trauma registry
(SweTrau). The registry includes data on vital signs, injuries,
interventions and patient information. Another database, the trauma care
quality database includes information relevant for MoM conferences. At
Karolinska University Hospital, patients who die during hospital stay
get their case sent to a MoM every time. However, for the rest of the
patients, two specialised nurses determine whether they are. Several
audit filters are set up to flag for patients with possible OFIs. Audit
filters may include low SBP, high RR, EMS scene time greater than 20
mins etc. Flagged patients are then reviewed by these two nurses who
decide if there is a possible OFI, in which case their case is sent to a
MoM conference for further review.

## **Outcome**

The outcome is the OFI delayed treatment. An OFI is any failure in care,
e.g. clinical judgement error, any preventable or potentially
preventable death, missed diagnosis, technical error and delayed
treatment as identified and decided my the MoM conference. The outcomes
we used were "long time until CT" and "long time until operation", which
will be referred to as delayed treatment from now on.

## **Participants**

The database, as previously mentioned, only includes patients 15 years
old and above, ISS \>9 and/or trauma team activation. Inclusion in the
study further required that the patients have been assessed for OFIs.
Patients were excluded if they were missing information in one of the
examined patient factors.

## **Variables**

The independent variables are gender, GCS on arrival, respiratory rate
on arrival, systolic blood pressure on arrival, number of injuries, ISS
and age. Of these variables, GCS, respiratory rate, systolic blood
pressure, number of injuries, ISS and age are used as continuous
variables in the analysis. Gender is used as a categorical variable.

For missing hospital values in GCS, RR and SBP, values collected from
EMT personnel before arrival to hospital were used. This was
accomplished by identifying missing data in the column for the hospital
value of the variable and, if available, replacing the row in the
dataset with the prehospital value. If a prehospital value is missing,
the patient is excluded from the study.

## **Data sources/measurement**

The data comes from the trauma registry and trauma care quality database
and is accessed through the program rstudio.

## **Bias**

Selection bias as this is a retrospective post hoc cohort study where
the outcome has occurred before the study began. This might have
affected how we chose variables, opting for variables with more
interesting results.

Misclassification bias as the conference may reach a consensus that is
wrong, alternatively that the two specialised nurses going through the
patients flagged for audit filters might misclassify patients or that
the audit filter misses a patient with ofi.

## **Study size**

The study size is all of the eligible patients that were treated at the
hospital between 2012 and 2022. Starting 2013 only a subgroup of
patients were screened for OFIs, but starting 2017 all patients that are
included in the database are screened for OFIs. Out of the 14,022
patients in the database, 7,378 were included in the study, with the
rest being excluded due to either not being screened for ofi or missing
values in the examined variables. 72 patients were deemed to have
received delayed treatment. However, only 70 patients with the
registered outcome were included in the study due to missing values
among two of the patients.

## **Statistical methods**

Multivariable and univariable logistic regression. A 5% significance
level and 95% confidence level was used. Rstudio was used to clean and
prepare the data from the database. The data was processed to have the
variables customised to fit into the logistical regression model, such
as making them into factors, renaming outcomes in the dataframe and
replacing or removing unknown values. The variables were either binary
or continuous. While writing the code, it was tested on scrambled data
so as to avoid selection bias as much as possible. When the code was
finished, it was run on the unscrambled data for a result.

Using the one in ten rule, 7 independent variables were used for the
logistical regression given that we have 70 patients with delayed
treatment. The one in ten rule states that for every variable examined,
10 events are needed in the group. In this case, this means that because
we have 70 patients with the examined outcome, we can have 7 variables.

The logistical regression was run through Rstudio using the glm command.
It was run several times to make an adjusted table and 7 unadjusted, one
for each variable. To make the tables, the package gtsummary was used to
visualise the adjusted and unadjusted regression individually, then
merging them together into one table.

This process was the only time generative AI was used in the writing
process. It was used to make snippets of code for a specific purpose,
revise and optimise code and find solutions when written code was not
functioning properly.

## **Ethical considerations**

An ethical problem that might be brought up is the anonymity of the
patients. In the registry, a patient's social security number is present
and could theoretically be used to identify the patient in question.
This will not be a problem though, as in the analysis, the social
security number is removed from the dataframe before the logistical
regression was run. Furthermore, the results will be presented with
patients grouped together, not individually. This will make it
impossible for anyone to identify patients. During the analysis, a VPN
was used to access the database to ensure that no data was leaked or
used for other purposes. Ethical permit is required and exists. The
ethical review number is 2021-02541 and 2021-03531.

# Results

Out of the 14,022 patients receiving trauma care at Karolinska
University Hospital, 5,710 patients were not screened for ofi, which
excludes them from the study and leaves 8,312 patients that have been
screened for ofi.

As we were running the logistic regression, additional patients were
excluded from the study due to missing data. Among the 8,312 patients
screened for ofi, 7,378 were included in the study. 934 patients were
excluded from the study due to missing values in one or more variables.
This left us with 70 patients in the delayed treatment group and 7,308
patients in the no delay to treatment group. Figure 1 below illustrates
the exclusion and inclusion of patients.

**Figure 1. The flowchart illustrates how patients were included and
excluded from the study, with what group they are in and the amount of
patients in that group illustrated in the boxes**

![](Rplot.png)

Figure 2 shows the characteristics of the group used in the logistic
regression. Most of the patients were men, 68%, with that being a bit
higher in the delayed treatment group, 74%. The mean value for the ISS
score was 10.18, with that value being higher in the outcome group, at
16.81. The mean age of the patients was 45.18 years old, with it being
around the same in the delayed treatment group at 45.50. The mean number
of injuries among the delayed treatment patients was at 6.33, which is
higher than in the overall group at 4.76. The mean SBP was 137.33, with
a standard deviation of 25.69. Mean respiratory rate was 18.59 with it
being slightly higher in the outcome population at 20.39. Lastly, mean
GCS was 13.93, being lower among patients with delayed treatment at
12.56.

**Figure 2. Characteristics of patients included in study. Shown is the
amount of participants in each group and overall with mean values and
standard deviations, except in gender, where the amount of participants
in each category is shown instead.**

```{r message = FALSE}
table1
```

**Abbreviations: N - number of patients, ISS - Injury Severity Score,
GCS - Glasgow Coma Scale.**

**a P-value shows fisher's test for gender and Welch t-test for the rest
of the variables.**

In the following table, figure 3, the results from the logistical
regression analysis is displayed.

The odds ratio is calculated both unadjusted and adjusted. A 95%
confidence interval was used with the p value shown in the table.

The analysis shows that some variables seem to have an association with
patients receiving delayed treatment. ISS and respiratory rate (RR) both
have a trend that patient with lower values seem to have a lower
incidence of delayed treatment, in both unadjusted (ISS, OR 1.04, 95% CI
1.02-1.05)(RR, OR 1.06, 95% CI 1.02-1.09) and adjusted analyses (ISS, OR
1.03, 95% CI 1.01-1.06)(RR, OR 1.04, 95% CI 1.01-1.08). GCS had the
opposite correlation, with higher GCS often having lower incidence of
delayed treatment, both unadjusted (OR 0.88, 95% CI 0.83-0.94) and
adjusted (OR 0.93, 95% CI 0.87-1.00). Number of injuries had the
association of a higher number of injuries having higher incidence of
delayed treatment unadjusted (OR 1.06, 95% CI 1.02-1.10), but that
association was not seen when adjusted for. The three remaining
variables, age, gender and SBP, were not shown to be linked to the
outcome.

**Figure 3. Results of logistic regression with odds ratio, 95%
confidence interval and p-value. Both unadjusted values and adjusted
values are shown.**

```{r message = FALSE}
result1
```

**Abbreviations: OR - Odds ratio, CI - Confidence interval, ISS - Injury
Severity Score, GCS - Glasgow Coma Scale.**

# **Discussion**

The aim was to determine the associations between patient level factors
and receiving delayed treatment in adult trauma patients. A logistic
regression on seven variables showed that higher ISS and RR and lower
GCS was significantly associated with patients receiving delayed
treatment.

This suggests that a heavily injured and less conscious patient is more
prone to receiving delayed treatment. Compared to a previous study
analysing the same data but with OFI as the observed outcome, this
result is similar to what they found[@Pat_factors]. As the mentioned
study suggests, this could be explained by that lower GCS as well as a
higher ISS indicates severe trauma. More injured patients often require
more complex management and a higher number of interventions, increasing
the risk of delayed treatment.

Respiratory rate had the correlation that a lower respiratory rate was
associated with lower rates of delayed treatment among the patients. As
the interval is \>10 to \<29 in the triage guideline, both too low and
too high respiratory rate is not desirable[@EHCP2015]. This could have a
similar explanation to ISS and GCS, that patients with tachypnea might
receive higher rates of delayed treatment because they are more severely
injured, therefore a higher risk of missing a possible injury. The
previous study also found this, but the correlation disappeared when
adjusted for, which said study hypothesised being due to its shared
variability with the ISS variable[@Pat_factors]. In this study, the same
correlation could be seen with NOI. A more severely injured patient will
score higher on both ISS and NOI. Therefore, our thought is that the
effect of NOI becomes insignificant when adjusting for ISS.

Not many studies regarding trauma and patient level factors have had
delayed treatment as their outcome, as many are analysing mortality.
Some similarities exist between which patient level factors predict the
outcome, with higher mortality being seen in patients with lower GCS and
higher ISS [@Mort_sammy][@Mort_hefny]. Advanced age has also been
identified as a contributory factor to increased mortality. However, it
is noteworthy that this study did not establish a significant
association between age and delayed treatment. Unlike ISS and GCS, which
are indicative of the severity of injuries and the immediate treatment
needs, advanced age is more likely to introduce comorbidities that
affect overall mortality but may not necessarily heighten the
probability of receiving delayed treatment[@Gioffre2018].

When discussing the outcome delay in treatment used in this study, it is
not a set time limit where if the wait exceeds it, it is considered
delayed treatment. The audit filters used at Karolinska University
Hospital flag for a wait for CT over 30 minutes and a wait for operation
over 60 minutes, which then gets reviewed by previously mentioned two
specialised nurses. However, if a patient has a longer wait, it does not
necessarily mean that there was a delay. 30 and 60 minutes are time
limits set to have a high sensitivity and low specificity in the audit
filters. Whether it is a delay is decided by the MoM conferences. It is
clear that the mean waiting time for CT and operation is higher than 30
and 60 minutes respectively when looking at the database. Using this
audit filter as the definition for delay would not mean anything, as the
majority of patients would be classified as having waited for too long.
Instead, the MoM conferences decide if the wait had any adverse effects
on the patients, such as an injury worsening during this time. It is
these cases that they label it as the \"too long until CT\" or \"Too
long until operation\". All delays in the database are delays that
adversely affected the patients.

#### **Strengths and Limitations**

In this study, one of the limitations was the number of patients having
the recorded outcome being quite small, leading us into having to limit
the number of variables we wanted to examine. Furthermore, the original
plan was to divide RR, SBP and GCS into revised trauma score (RTS)
categories from 1-4, since this would have allowed for a more detailed
analysis. However, the sample size was too small in the outcome group
which led us to resort to making these variables continuous. We
considered expanding the outcome group to include more OFI, but opted
for the solution used in the study instead as we felt that including
more OFI would lead to less clinically applicable results.

The problem mentioned above with too few patients in the outcome group
made the study into a post hoc analysis. We were not aware of the number
of patients who received delayed treatment when we handled the scrambled
data as we wanted to minimise selection bias. When we ran the code on
the real data it became clear that the sample size was too small for our
approach. Therefore, we had to change the variables after having seen
the real data, making it a post hoc analysis. This could in theory open
up for selection bias, as we might have chosen variables which give us
results we are looking for. We did our best to remove variables at
random, but it is possible that we subconsciously removed variables that
would change the result in a way benefiting us. The best way to have
gone about this would have been to calculate the amount of people who
received delayed treatment before choosing the variables. Then, choosing
an amount of variables that would work with the amount of patients. This
would have avoided making it post hoc and minimised selection bias.

On discussing possible biases, misclassification bias might also be
present in this study, though it is something that we cannot control in
this study. As previously mentioned, the possible existence of an OFI is
decided during the MoM conferences. However, they might not always be
correct in their judgement and occasionally misclassify, leading to
patients being in the wrong group in the analysis. Also, some patients
might have been missed even before reaching the MoM conference. The
patients included in the conference are selected through audit filters
and two specialised nurses, and patients might be missed in this stage.
This bias should not be big enough to have a significant impact on the
results, but can be kept in mind.

#### **Clinical applications**

Analysing the results and its clinical applications, the results can be
interpreted as that it is mostly injured patients with affected
consciousness that receive delayed treatment. The clinical applications
of this are that clinicians need to do more to avoid this delay for this
patient group. All patients in the outcome group are patients that
experienced adverse effects because of their delayed treatment. Some
might include injuries being missed or not being treated, leading to
unnecessary suffering for the patient. Clinicians being aware of the
patient level factors at risk for delayed treatment might help with
their decision making when caring for these patients.

It is important to highlight the importance of adequate diagnostic
imaging when the patient is stabilised, as a substantial part of the
patients in the outcome group were patients who waited too long for
diagnostic imaging. To lower this risk, it occurs that patients are
stabilised in the operating room (OR), gets taken to the CT then back to
the OR. This is done to minimise the risk of missing an injury,
especially in severely injured patients where it can be difficult to
identify all injuries without diagnostic imaging. When clinicians have
access to these instruments they should be used whenever deemed
necessary, as studies suggest it has a positive effect on patient
outcome [@Hajibandeh_2014].

#### Equity

The results of the study are expected as previous similar studies have
come to similar conclusions. A goal to strive towards in trauma care
through research and improvement programs is that no patient level
factors should be linked with a higher risk of delayed treatment. The
variables found in this study to be linked with delayed treatment are
patient factors that arise from the experienced trauma, not previously
existing in the patient such as age or gender. This implies that
patients are not systematically discriminated against based on their
gender or age, which is a positive sign seen from a health equity point
of view.

#### **Generalisability**

Discussing the generalisability of this study, we included all the
trauma patients treated and Karolinska University Hospital and excluded
40.7% of that population based on that they had not been screened for
OFI. Additional patients were excluded because of missing data, but no
distinct groups except the not screened patients were removed. On that
basis, this population should be a general representation of trauma
patients at the equivalent of a level 1 trauma center, making this study
generalisable. While the population only represents Karolinska
university hospital, the conclusion should be applicable to the whole
trauma population.

#### **Future studies**

To further expand the knowledge on what patient level factors are
associated with delayed treatment, future studies should aim for a
larger sample size to be able to analyse more variables, such as highest
hospital care level, intubation status or first emergency procedure
performed. To be able to do this, more patients in the outcome group
would be needed. This can be done by conducting a study which includes
the registries of several hospitals.

A question that arose during this study was what factors are associated
with other OFI, such as missed injury, communication errors or faulty
triage, and if they are similar to the ones found in this study. Future
studies could use the same data but use another OFI as the outcome to
compare how variables associated with each OFI differ.

## **Conclusion**

Based on this single trauma center, the results suggest that it is
mostly more injured and patients with affected consciousness that
receive delayed treatment. Clinicians should be aware of this as they
can take steps to lower that risk for this patient group. Further
studies need to be conducted on this subject to further examine the link
between delayed treatment and patient level factors.

# **Acknowledgements**

I would like to extend my gratitude to my supervisor Martin Gerdin and
my two co-supervisors Jonatan Attergrim and Kelvin Szolnoky for always
being quick with providing valuable help and feedback during all stages
of the writing process as well as guidance during the data analysis part
of the project.